---
title: "Clustering"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tibble)
library(lubridate)
library(GGally)
library(cluster)
```

## Task Implementation
Done some research and kmeans is apparently quite naive - working on
an implementation with pam (and pamk). Also need to work on how we
choose the number of clusters.

```{r data clean, message=FALSE}

spotify.clustering <- read_csv("inst/edited_spotify.csv")

clean_data <- spotify.clustering %>%
  mutate(AlbumReleaseDate = parse_date_time(AlbumReleaseDate, orders = c("y", "ym","ymd"))) %>%
  #Old-school grepl method
  mutate(Artist = ifelse(grepl("Beyonc*", Artist), 'Beyonce', Artist)) %>%
  #Tidyverse str_detect method
  mutate(Artist = ifelse(Artist %>% 
                           str_detect("Janelle Mon*"), 'Janelle Monae', Artist)) %>%
  mutate(Artist = ifelse(AlbumBestChartPosition %>% 
                           str_detect("#N/A"), 0, Artist)) %>%
  na.omit() %>%
  mutate(id = row_number()) %>%
  mutate(id = as.character(id))
sapply(data, class)

aggr(clean_data) # checks for missing data

test_data <- subset(clean_data, ((AlbumName == "A Girl Called Dusty") | 
                                 (AlbumName == "Action!") |
                                 (AlbumName == "Selling England By The Pound") |
                                 (AlbumName == "Carpenters") | 
                                 (AlbumName == "Ride On") |
                                 (AlbumName == "Autoamerican") |
                                 (AlbumName == "Selected Ambient Works 85-92") |    
                                 (AlbumName == "Different Class") |
                                 (AlbumName == "O") |
                                 (AlbumName == "The Elder Scrolls IV: Oblivion: Original Game Soundtrack") |
                                 (AlbumName == "AM") |
                                 (AlbumName == "An Awesome Wave")))

training_data <- subset(clean_data, ((AlbumName != "A Girl Called Dusty") & 
                                     (AlbumName != "Action!") &
                                     (AlbumName != "Selling England By The Pound") &
                                     (AlbumName != "Carpenters") & 
                                     (AlbumName != "Ride On") &
                                     (AlbumName != "Autoamerican") &
                                     (AlbumName != "Selected Ambient Works 85-92") &    
                                     (AlbumName != "Different Class") &
                                     (AlbumName != "O") &
                                     (AlbumName != "The Elder Scrolls IV: Oblivion: Original Game Soundtrack") &
                                     (AlbumName != "AM") &
                                     (AlbumName != "An Awesome Wave")))

training_data <- training_data[c("TrackDuration", "TrackDanceability",
                                 "TrackEnergy", "TrackKey", "TrackLoudness",
                                 "TrackSpeechiness", "TrackAcousticness",
                                 "TrackInstrumentalness", "TrackLiveness", "TrackValence",
                                 "TrackTempo")]

```


Defines the clustering of the training data
```{r clustering}
cluster_1 <- data %>%
  select(colnames(data[15:27]), -TrackMode) %>%
  kmeans(50,100,1)
```

```{r getter_functions}
get_knn <- function(){
  
}

get_row <- function(song_name){
  line_number <- data$TrackName %>%
    str_detect(song_name)
    
  return(min(which(line_number)))
}

get_cluster_rows <- function(cluster_number){
  clusters %>%
    filter(value == cluster_number) %>%
    select(id) 
  #Need to remove song currently listening to
}

#Object of type class NULL error...hmmm when trying to use dplyr slice but works with base :O
get_cluster <- function(row_number){
  cluster_1$cluster[row_number]
}

random_picker <- function(cluster_number){
new_song_rn <- get_cluster_rows(cluster_number) %>%
  #So here instead of randomly picking 1, would be good to
  #call another function based on the genre and say popularity?
    sample_n(1)
data %>%
  filter(id == as.character(new_song_rn)) %>%
  select(TrackName)
}

#sudo - subset data based on cluster and remove existing song -> then pick one of remaining songs.
pick_new_song <- function(song_name){
  cluster_number <- get_cluster(get_row(song_name))
  random_picker(cluster_number)
}
pick_new_song("Sandy's Holiday")

#can look at following: cluster, centers, withins, size
#Could simply link each song to a cluster, pick nother song from that cluster based on a probabilty linked to popularity of the song and maybe an artist variables as well, as in pick same artist with higher prob if it is in the cluster
```

## Gabriels's Code
 expenential -0 is the maybe the way to go (scaled)
 database of thousands of songs - maybe be wary of outliers and such not
 simply twidderlerably parameters, raising questions so our model +how could we improve
 scaleability
 MEAN - PROVIDE SOME INFORMATION BUT LOSES ALOT OF DATAPOINTS. 
 WHY RATHER THAN ACTUAL METHOD, TALK ABOUT METHODOLOGY - MORE IMPORTANT
 think about aim of the recommendation - element of exploration - objective function
 WHAT WE FOUND OUT WHAT WE CONCLUDE AND WHAT WE RECOMMEND AND BUILD A REPORT ABOUT THAT AND SJUSTIFY WHAT WE HAVE DONE - WEEK 1 STAT INVESTICATION CYCLE
can look at following: cluster, centers, withins, size
Could simply link each song to a cluster, pick nother song from that cluster based on a probabilty linked to popularity of the song and maybe an artist variables as well, as in pick same artist with higher prob if it is in the cluster

